Efficient pedigree recording for fast population genetics simulation

### 2018-01-11 forwards_paper.tex

\krt{Are we doing a methods section?}

\subsection{Requirements for the \cpp{} implementation}

\begin{enumerate}
    \item \fwdpy{} 0.13.a0
    \item \fwdpp{} 0.5.7
    \item GNU Scientific Library 1.16
    \item pybind11 2.2.1
    \item GCC 4.8.5
    \item \msprime{} master branch \jda{which commit hash or short version number
		    from pip freeze or {\tt msprime.\_\_version\_\_}?}
\end{enumerate}

All of the above installed via conda, except for msprime, which must be installed from GitHub.

\subsection{Requirements for the \simupop{} implementation}

The simulations are implemented the Python package \ftprime{} 0.0.6rc0
which is available at \url{https://www.github.com/ashander/ftprime}.
Its requirements are encoded in an \texttt{environment.yml} file, compatible
with conda, but \msprime{} must be updated to the version listed below via pip
from GitHub \jda{TODO add command}.
In brief the requirements are:

\begin{enumerate}
    \item \msprime{} \jda{TODO add version from KRT's dev machine}
    \item \simupop{} 1.1.8.3
    \item python 3.5
\end{enumerate}

>>>>

\subsection*{The \cpp{} implementation}

\paragraph{Requirements.}
The \cpp{} implementation uses several packages installed via conda,
except for \msprime{}, which must be installed from GitHub.

\begin{enumerate}
    \item \fwdpy{} 0.13.a0
    \item \fwdpp{} 0.5.7
    \item GNU Scientific Library 1.16
    \item pybind11 2.2.1
    \item GCC 4.8.5
    \item \msprime{} 0.5.0 \jda{Changed these but the actual sims were not run
		    with this version...}
\end{enumerate}

\subsection*{The \simupop{} implementation}

We implemented simulations in a Python script, usable as a command-line utility,
taking input parameters for population size $N$ and scaled recombination rate $\rho$.
As described in 'Simulation benchmarks', we set scaled mutation rate $\theta=\rho$,
we used simplify interval of 1000 and simulated for $T=10N$ generations in each
simulation, introducing deleterious mutations at a rate of $\rho/100$ per
generation.
Because \simupop{} requires explicit locations for loci, we set the number of
loci $n=\rho / (4Nr)$ and marked
$k={\rm round}(\mu / (100 (\mu + \mu / 100))) = 0$ of these as selected loci.
\jda{unfortunately probably what we did...}
The selected loci were chosen as a random sample of all the loci.

We coupled \simupop{} to \msprime{} using a Python class \texttt{RecombCollector} that 1)
tracks time, and 2) collects and parses recombination information provided by
the \texttt{Recombinator} class of \simupop{}, adding them to a \tnodetable{} and
\tedgetable{} upon which simplification is performed every 1000 generations (using
\texttt{msprime.sort\_tables()} and \texttt{msprime.simplify\_tables}).
The tables are initialized using the \texttt{TreeSequence} returned from
\texttt{msprime.simulate($2N$, Ne=$N$, recombination\_rate=$r/2$, length=$n$)}.
The \texttt{simuPOP.Population()} is initialized and tagged with individual IDs,
a mapping of these IDs to the \tnodetable{} IDs is passed to
an instance of \texttt{RecombCollector} and used internally to update the tables.
The \texttt{simuPOP.Population()} is then evolved.
Before mating: selected mutations are created (using
\texttt{simuPOP.SNPMutator}) and multiplicative selection occurs (using
\texttt{simuPOP.PyMlSelector}) only on the selected loci;
the internal time of the \texttt{RecombCollector} instance is updated.
During mating: male and female parents are drawn randomly, recombination data is
output by \simupop{} to the \texttt{RecombCollector} instance.
After mating: nothing happens 999/1000 generations, every 1000 generations the tables are
simplifed.

\texttt{RecombCollector} uses the Python class \texttt{ARGrecorder} internally
to add data to the tables and perform simplification.
This class contains the \tnodetable{} and \tedgetable{} and also the mapping from
(haploid) ids to ids in the tables.
It also provides methods (\texttt{add\_individual} and \texttt{add\_record})
to append data to the tables.
The only functions of
\texttt{RecombCollector} not delegated to this internal class are computing
haploid IDs from \simupop{}'s diploid IDs and parsing recomination data into
data suitible for use with use with \texttt{add\_individual} and
\texttt{add\_record}; these functions are used to stream data into the tables as
it is received by an instance of \texttt{RecombCollector}.

\paragraph{Requirements.}
The simulations as implemented depend on the class \texttt{RecombCollector} provided by
the Python package \ftprime{} 0.0.6rc0,
available at \url{https://www.github.com/ashander/ftprime}.
Its requirements are encoded in an \texttt{environment.yml} file, compatible
with conda, but \msprime{} must be updated to the version listed below via pip
from GitHub \jda{TODO add command?}.

In brief the requirements are:

\begin{enumerate}
    \item \msprime{} 0.5.0
    \item \simupop{} 1.1.8.3
    \item python 3.5
\end{enumerate}
